---
title: "Food at College"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

```{r, warning=FALSE, message=FALSE}
# From your file menu (at the top of R-Studio) select:
# "Session -> Set working directory -> To source file location"
# Then play this chunk to get the data into R.
library(mosaic)
library(car)
library(DT)
library(pander)
library(readr)
library(tidyverse)
food <- read_csv("../../Data/food.csv") #food.csv is in the Data folder...
View(food)
```


----

<!-- Instructions:

 Use the Food dataset and a Kruskal-Wallis Test(s) to answer an interesting question(s) that you come up with.
 
 About the dataset:
 
  * GPA - numerical, self-reported current college GPA

  
  * weight - numerical, self-reported weight in lbs


  * gender - "Female", "Male"


  * breakfast - "Cereal", "Donut" 
      The participants were shown a picture of "Cereal" and a "Donut" and asked which one of these pictures they associate with the word "breakfast."


  * calories_chicken - guessing calories in chicken piadina from the options of:
     1 - 265 
     2 - 430 
     3 - 610 
     4 - 720
    (the variable shows the actual number of calories participants selected)


   * calories_day - Importance of consuming calories per day
     1 - i dont know how many calories i should consume 
     2 - it is not at all important 
     3 - it is moderately important 
     4 - it is very important


   * comfort_food - List 3-5 comfort foods that come to mind. 
     Open ended 


   * comfort_food_reasons - What are some of the reasons that make you eat comfort food?      (i.e., anger, sadness, happiness, boredom, etc) - list up to three
     Open ended


   * comfort_food_reasons_coded
     1 - stress
     2 - boredom
     3 - depression/sadness
     4 - hunger
     5 - laziness
     6 - cold weather
     7 - happiness 
     8 - watching tv
     9 - none 


    * cook - how often do you cook?
     1 - Every day 
     2 - A couple of times a week 
     3 - Whenever I can, but that is not very often  
     4 - I only help a little during holidays 
     5 - Never, I really do not know my way around a kitchen


    * cuisine - what type of cuisine did you eat growing up?
     1 - American
     2 - Mexican.Spanish
     3 - Korean/Asian
     4 - Indian
     5 - American inspired international dishes
     6 - other


    * diet_current - describe your current diet
     open ended

	
    * diet_current_coded
     1 - healthy/balanced/moderated/
     2 - unhealthy/cheap/too much/random/
     3 - the same thing over and over
     4 - unclear


    * drink - which picture do you associate with the word “drink”?
     1 - orange juice
     2 - soda 


    * eating_changes  - Describe your eating changes since the moment you got into college?
     Open ended 


    * eating_changes_coded
     1 - worse
     2 - better
     3 - the same
     4 - unclear


    * eating_changes_coded1
     1 - eat faster
     2 - bigger quantity
     3 - worse quality 
     4 - same food
     5 - healthier
     6 - unclear
     7 - drink coffee 
     8 - less food
     9 - more sweets
     10 - timing 
     11 - more carbs or snacking
     12 - drink more water
     13 - more variety


    * eating_out - frequency of eating out in a typical week 
     1 - Never 
     2 - 1-2 times 
     3 - 2-3 times 
     4 - 3-5 times 
     5 - every day


    * employment - do you work? 
     1 - yes full time 
     2 - yes part time 
     3 - no
     4  - other


    * ethnic_food - How likely to eat ethnic food 
     1 - very unlikely 
     2 - unlikely 
     3 - neutral 
     4 - likely 
     5 - very likely 


    * exercise - how often do you exercise in a regular week?
     1 - Everyday 
     2 - Twice or three times per week 
     3 - Once a week
     4 - Sometimes 
     5 - Never


    * father_education - 
     1 - less than high school 
     2 - high school degree 
     3 - some college degree 
     4 - college degree 
     5 - graduate degree 


    * father_profession - what is your father profession?
     Open ended


    * fav_cuisine - What is your favorite cuisine?
     Open ended


    * fav_cuisine_coded
     0-none
     1 - Italian/French/greek
     2 - Spanish/mexican
     3 - Arabic/Turkish
     4 - asian/chineses/thai/nepal
     5 - American
     6 - African 
     7 - Jamaican
     8 - indian


    * fav_food - was your favorite food cooked at home or store bought? 
     1 - cooked at home 
     2 - store bought 
     3 - both bought at store and cooked at home


    * food_childhood - what was your favorite childhood food?
     Open ended



    * fries - which of these pictures you associate with word fries? 
     1 - Mcdonald’s fries
     2 - home fries

 
    * fruit_day - How likely to eat fruit in a regular day 
     1 - very unlikely 
     2 - unlikely 
     3 - neutral 
     4 - likely 
     5 - very likely 


    * grade_level – 
     1 - freshman 
     2 -Sophomore 
     3 - Junior 
     4 - Senior
     
      
    * greek_food - How likely to eat greek food when available?
     1 - very unlikely 
     2 - unlikely 
     3 - neutral 
     4 - likely 
     5 - very likely 
     
      
    * healthy_feel – how likely are you to agree with the following statement: “I feel very healthy!” ?
      1 to 10 where 1 is strongly agree and 10 is strongly disagree - scale
      
      
    * healthy_meal – what is a healthy meal? Describe in 2-3 sentences.
      Open ended
      
      
    * ideal_diet – describe your ideal diet in 2-3 sentences
      Open ended
      
      
    * Ideal_diet_coded
      1 – portion control
      2 – adding veggies/eating healthier food/adding fruit	
      3 – balance
      4 – less sugar
      5 – home cooked/organic
      6 – current diet
      7 – more protein
      8 – unclear
      
      
     * income
      1 - less than $15,000 
      2 - $15,001 to $30,000 
      3 - $30,001 to $50,000 
      4 - $50,001 to $70,000 
      5 - $70,001 to $100,000 
      6 - higher than $100,000
      
     * indian_food – how likely are you to eat indian food when available
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4 - likely 
      5 - very likely 
      
      
     * Italian_food – how likely are you to eat Italian food when available?
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4 - likely 
      5 - very likely 
      
      
     * life_rewarding – how likely are you to agree with the following statement: “I feel life is very rewarding!” ?
      1 to 10 where 1 is strongly agree and 10 is strongly disagree - scale
      
      
     * marital_status
      1 -Single 
      2 - In a relationship 
      3 - Cohabiting 
      4 - Married 
      5 - Divorced 
      6 - Widowed
      
      
     * meals_dinner_friend – What would you serve to a friend for dinner?
      Open ended
      
      
     * mothers_education 
      1 - less than high school 
      2 - high school degree 
      3 - some college degree 
      4 - college degree 
      5 - graduate degree
      
      
     * mothers_profession – what is your mother’s profession? 
      
      
     * nutritional_check - checking nutritional values frequency 
      1 - never 
      2 - on certain products only 
      3 - very rarely 
      4 - on most products 
      5 - on everything
      
      
     * on_off_campus – living situation
      1 - On campus 
      2 - Rent out of campus 
      3 - Live with my parents and commute 
      4 - Own my own house
      
      
     * parents_cook - Approximately how many days a week did your parents cook? 
      1 - Almost everyday 
      2 - 2-3 times a week 
      3 - 1-2 times a week 
      4 - on holidays only 
      5 - never
      
      
     * pay_meal_out - How much would you pay for meal out? 
      1 - up to $5.00 
      2 - $5.01 to $10.00 
      3 - $10.01 to $20.00 
      4 - $20.01 to $30.00 
      5 - $30.01 to $40.00 
      6 - more than $40.01 
      
      
     * Persian_food - How likely to eat persian food when available?
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4 - likely 
      5 - very likely 
      
      
     * self_perception_weight - self perception of weight 
      6 - i dont think myself in these terms 
      5 - overweight 
      4 - slightly overweight 
      3 - just right 
      2 - very fit 
      1 - slim 
      
      
     * Which of the two pictures you associate with the word soup?
      1 – veggie soup
      2 – creamy soup
      
      
     * sports - sports – do you do any sporting activity?
      1 - Yes 
      2 - No 
      99 – no answer
      
      
     * thai_food - How likely to eat thai food when available?
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4 - likely 
      5 - very likely 
      
      
     * tortilla_calories - guessing calories in a burrito sandwhich from Chipolte?
      1 - 580 
      2 - 725 
      3 - 940 
      4 - 1165
      
     * turkey_calories - Can you guess how many calories are in the foods shown below?
     (Panera Bread Roasted Turkey and Avocado BLT)
      1 - 345 
      2 - 500 
      3 - 690 
      4 - 850
      
      
     * type_sports – what type of sports are you involved?
      Open-ended
      
     * veggies_day - How likely to eat veggies in a day? 
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4- likely 
      5 - very likely
      
     * vitamins – do you take any supplements or vitamins?
      1 – yes
      2 – no
      
     * waffle_calories - guessing calories in waffle potato sandwhich 
      1 - 575 
      2 - 760 
      3 - 900 
      4 - 1315


Note this dataset and description come from: © 2020 Kaggle Inc for further details visit:

[Food choices on Kaggle](https://www.kaggle.com/borapajo/food-choices?select=food_coded.csv)


If you want to give your critiquers some ideas about what you have questions on and would like help with, place those details here.-->




<!--End your comments to your critiquers before this line.-->

----

<!--Begin writing your Analysis after this point.-->

## Background

'Food' is a dataset containing information taken from surveys of college students. From 3 categories in our dataset titled 'life_rewarding', 'mother_education', and 'father_education' we can ask the question, "Does the education of somebody's parents have an impact on their happiness in life?" While there are subjective elements in reporting, we can see what information we can draw from a statistical analysis.

## Data

<<<<<<< HEAD
To answer this question, we need to modify the dataset. We will clean the data set, then sum the parent's education (mother_education + father_education), and then split the data according to whether the total is over 6 or under 6. We choose 6 because the scales of mother_education and father_education are 1-5, so the lowest data value in the sum is 2, and the maximum is 10, with the halfway value being 6. The scale does not reflect an objective metric of parental education, but is only a perceived scale. In that same vein, life_rewarding is also on a subjective scale from 1-10. We can still see if there are any patterns in the data.
=======
To answer this question, we need to modify the dataset. We will clean the data set, then sum the parent's eduction (mother_education + father_education), and then split the data according to whether the total is over 6 or under 6. We choose 6 because the scales are 1-5, so the lowest data value in the sum is 2, and the maximum is 10.


!! what does the difference in parent education and life rewarding mean? give background
!! What do the variables mean? in the chart?
!! add information, explain what it is
!! describe what's going on
>>>>>>> ff835129b138e6c7ad4bae421438fee2f16ec3e6

```{r}

data_cleaned <- subset(food, !is.na(father_education) & !is.na(mother_education) & !is.na(life_rewarding))


life_r_ed <- data_cleaned %>% mutate(
  sum_ed = mother_education + father_education,
<<<<<<< HEAD
  parent_ed_6 = case_when(
    sum_ed > 6 ~ "Over 6",
    sum_ed <= 6 ~ "Under 6"
  ))



# Count occurrences where sum_ed is "Under 6" and parent_ed_6 is "Under 6"
# over_6_count <- sum(life_r_ed$parent_ed_6 == "Over 6")

# Count occurrences where sum_ed is "Under 6"
# under_6_count <- sum(life_r_ed$parent_ed_6 == "Under 6")

# Print the counts
# print(over_6_count)
# print(under_6_count)


```


## Test and Hypothesis

We will be running a Kruskal Wallis test on the data. Typically the Kruskal Wallis test is ideal when the data is under n = 30 or is non normally distributed. Our n is higher than 30 and somwhat normally distributed, but we will still run the test.

We will set our p value at .05.

The null hypothesis is that there is no difference in the distributions between 'life_rewarding' reports of people whose parents have a total education greater than 6, and those whose parents have a total education less than or equal to 6, and the alternative hypothesis is that there is a difference in those distributions, or that they are from different distributions.

H0:All samples are from the same distribution.
Ha:At least one sample's distribution is stochastically different.

\[
H_0: \mu_{\text{life_rewarding}_{(\text{parent_education} > 6)}} = \mu_{\text{life_rewarding}_{(\text{parent_education} \leq 6)}}
\]

\[
H_a: \mu_{\text{life_rewarding}_{(\text{parent_education} > 6)}} \neq \mu_{\text{life_rewarding}_{(\text{parent_education} \leq 6)}}
\]


## Boxplot

Let's take a look at a boxplot of life_rewarding responses split by the parent's education.

```{r}

boxplot(life_rewarding ~ parent_ed_6,data = life_r_ed, col=c("red","blue"))
```

We can visually see a difference here, with Under 6 having a median life_rewarding response of 4 and over 6 having a median of 6.

## qqPlot

Let's take a look at the normality of the data using a qqplot.

```{r}
qqPlot(life_rewarding ~ parent_ed_6, data = life_r_ed)
=======
  over6 = case_when(
    sum_ed > 6 ~ 1,
    sum_ed <= 6 ~ 0
  ))

```


## Hypothesis

The null hypothesis is that there is no difference in the distributions between 'life_rewarding' reports of people whose parents have a total education greater than 6, and those whose parents have a total education less than or equal to 6, and the alternative hypothesis is that there is a difference in those distributions, or that they are from different distributions.

H0:All samples are from the same distribution.
Ha:At least one sample's distribution is stochastically different.

\[
H_0: \mu_{\text{life_rewarding}_{(\text{parent_education} > 6)}} = \mu_{\text{life_rewarding}_{(\text{parent_education} \leq 6)}}
\]

\[
H_a: \mu_{\text{life_rewarding}_{(\text{parent_education} > 6)}} \neq \mu_{\text{life_rewarding}_{(\text{parent_education} \leq 6)}}
\]


## Food
>>>>>>> ff835129b138e6c7ad4bae421438fee2f16ec3e6



<<<<<<< HEAD
We can see that the data is mostly normal, except for the bottom and top for both groups. Note that there is an inherent skew because life_rewarding is a step-wise scale and not continuous, which forces values that could otherwise be in between numbers up or down and creates a staircase effect.

## Kruskal Test Results
=======
```{r}

test <- kruskal.test(life_rewarding ~ over6, data = life_r_ed)


boxplot(life_rewarding ~ over6,data = life_r_ed, col=c("red","blue"))

qqPlot(life_rewarding ~ over6, data = life_r_ed)
# qqPlot(life_rewarding ~ sum_ed, data = life_r_ed)

test

summary(test)



```

>>>>>>> ff835129b138e6c7ad4bae421438fee2f16ec3e6

Let's take 

```{r}

test <- kruskal.test(life_rewarding ~ parent_ed_6, data = life_r_ed)
test

```


Our p value is .2001 which means that there is not enough evidence to suggest that the distributions are different. Our chi-squared value was 1.6419 with df = 1.

## Conclusion

Although there is a visual difference in the box plot, the Kruskal Wallis test does not show that there is a statistically significant difference. This would suggest, at least from a subjective standpoint, that people's perceptions of how rewarding their life is are not from different distributions based on their perception of their parent's education levels.





