---
title: "Car Prices"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

<br>

```{r, include=FALSE}
library(dplyr)
library(tidyverse)
library(plotly)
library(ggplot2)
library(pander)
library(car)
library(lattice)

# Play this chunk to load the data.
CarPrices <- read.csv("../../Data/CarPrices.csv", header=TRUE, quote="\"")

Deville <- filter(CarPrices,CarPrices$Make == "Cadillac" & CarPrices$Model == "Deville")
```

## Background

CarPrices is a data set that contains 80 samples of Cadillac cars. We will take a look specifically at Price as a factor of Mileage for the Deville model, and take a look at how the kind of Trim style the Deville model has greatly influences the price.

## Model

Below is the model for simple linear regression

$Y_i = \beta_0 + \beta_1 X_{1i} + \beta_2 X_{2i} + \epsilon_i$





```{r}
# Assuming 'Deville' is your dataframe containing the data

deville_lm <- lm(Price ~ Mileage, data = Deville)

b <- coef(deville_lm)

p <- plot_ly(data = Deville, x = ~Mileage, y = ~Price, type = "scatter", mode = "markers", 
             color = ~Trim,
             text = ~paste("Trim: ", Trim)) %>%
  layout(title = "Price vs Mileage",
         xaxis = list(title = "Mileage"),
         yaxis = list(title = "Price"))

# Add regression line to the plot
p <- add_trace(p, x = Deville$Mileage, y = b[1] + b[2]*Deville$Mileage, 
               type = "scatter", mode = "lines", line = list(color = "green"))

# Print the plot
p

```

```{r}

# Set up a 1x2 grid for plots
par(mfrow=c(1,3))
plot(deville_lm, which=1)
qqPlot(deville_lm$residuals, id=FALSE)
plot(deville_lm$residuals)

summary(deville_lm)

```

Adjusted R-squared:  0.3475

## Revised Model



Below is the equation for a two lines linear regression model

$\underbrace{Y_i}_\text{Price} = \underbrace{β0 + β1X1i + β2X2i}_\text{ E(Yi)} + ϵ_i$


\[ X_{2i} = \begin{cases}
1 & \text{if Trim = DHS Sedan 4D or DTS Sedan 4D"} \\
0 & \text{if Trim = Sedan 4D}
\end{cases} \]

```{r}

# Create the scatter plot using plotly, coloring points based on the "Trim" variable


p <- plot_ly(data = Deville, x = ~Mileage, y = ~Price, type = "scatter", mode = "markers", 
        color = ~Trim,
        text = ~paste("Trim: ", Trim)) %>%
  layout(title = "Price vs Mileage",
         xaxis = list(title = "Price"),
         yaxis = list(title = "Mileage"))

b <- coef(deville_lm)

p <- add_trace(p, x = Deville$Mileage, y = b[1] + b[2]*Deville$Mileage, 
               type = "scatter", mode = "lines", line = list(color = "skyblue"))

p

# Find a variable for cadillac deville that you can add, that if you add, makes for a better fit.
# Try 2 lines, see what splits the values. Look at r2 values,
# Do just mileage
# When you add the variable, you will see the r2 jump.


```

split based on trim
copy equation except for last change in slope term from statstics notebook
do 3 graphs for new lines with regression model

```{r}
Deville <- Deville %>%
  mutate(
    Trim_Case = case_when(
      Trim %in% c("DHS Sedan 4D", "DTS Sedan 4D") ~ 1,
      Trim == "Sedan 4D" ~ 0
    )
  )


# Fit linear models
lm_trim <- lm(Price ~ Mileage + Trim_Case, data = Deville)

# Obtain fitted values
# fitted_values1 <- predict(lm_dts_dth)

# Get coefficients
bd <- coef(lm_trim) 

# Create scatter plot
p <- plot_ly(data = Deville, x = ~Mileage, y = ~Price, type = "scatter", mode = "markers", 
             color = ~Trim,
             text = ~paste("Trim: ", Trim)) %>%
  layout(title = "Price vs Mileage",
         xaxis = list(title = "Mileage"),
         yaxis = list(title = "Price"))

# Add regression lines
p <- add_trace(p, x = Deville$Mileage, y = bd[1] + bd[2]*Deville$Mileage, 
               type = "scatter", mode = "lines", line = list(color = "lightblue"))

p <- add_trace(p, x = Deville$Mileage, y = (bd[1] + bd[3]) + bd[2]*Deville$Mileage, 
               type = "scatter", mode = "lines", line = list(color = "pink"))

p  # Print the plot

```

```{r}

par(mfrow=c(1,3))
plot(lm_trim, which=1)
qqPlot(lm_trim, id=FALSE)
plot(lm_trim$residuals)

summary(lm_trim)

```



R squared: 0.9572 

