---
title: "Wk 10 Class Notes - Monday (Visualizing complicated data)"
execute:
  keep-md: TRUE
  df-print: paged
  warning: false
format:
  html:
    theme: cerulean
    code-fold: true
    code-line-numbers: true
date: "2024-06-24"
---

```{r}
library(tidyverse)
library(mosaic)
View(Births78)
```

```{r}
B78 <- Births78 %>% mutate(weekend = ifelse(wday %in% c("Sat", "Sun"), 1,0))
pairs(B78, panel=panel.smooth, col=as.factor(B78$weekend))
```
```{r}

```



```{r}
ggplot(B78,aes(x=day_of_year,y=births,col=weekend)) +
  geom_point() +
  facet_wrap(interaction(B78$weekend)) +
  geom_smooth(method="loess", span= 0.5)
```
```{r}

```

Expert opinion - crazy good data. No, tell them what they want to hear. Good idea, but don't trust them just because they say you should...

How do we attack the data? esp if we can't even see the pairs plant?

If I can't see it, I don't believe it, when it comes to data.

In God we trust, all others bring data.

"What am I looking at?"
Try to see what your model is doing according to the data. Draw it.

When you push into high dimensional datasets, unless you can visualize it, how do you move forwards? You're gonna be lost.

Having sale price made the model prediction worse?

R^2 vs Interpretation vs Visualization
The higher R^2 can lose interpretation and visualization
Interpretable and visualizable models could be less complicated.

Kills interpretation:
- Transformations
- Anything that's not a line?
- Log is interpretable
- Pretty much anything else is not
