---
title: "Residuals and Sums of Squares"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
    toc: true
    toc_float: true
---



# Week 2 Mobius - Residuals

```{r}
pacman::p_load(tidyverse,dplyr,car,pander)
```

```{r}
View(Orange)
```



```{r}

mylm <- lm(circumference ~ age, data=Orange)
summary(calm)
```


# b0/intercept, b1/slope

```{r}
y <- 17.39965 + 0.106770

y <- 17.39965 + 0.106770 * 3 * 365
y
```


```{r}
plot(circumference ~ age, data=Orange,col="orangered",pch=16)
abline(calm,col="blue")
```






# Y is the actual Y value of the data set - circumference

```{r}

sum((Y - mylm$fit)^2 )
sum((Orange$circumference - mylm$fit)^2 )

sum((mylm$fit - mean(Y))^2 )
sum((mylm$fit - mean(Orange$circumference))^2 )

sum( (Y - mean(Y))^2 )
sum((Orange$circumference - mean(Y))^2 )
```



# Correlation

```{r}
cor(Orange$circumference,Orange$age)

clm1 <- lm(mpg ~ wt, data = mtcars)
clm2 <- lm(mpg ~ cyl, data = mtcars)
clm3 <- lm(mpg ~ hp, data = mtcars)
```




# Set up the plotting area to have 1 row and 3 columns
```{r}
par(mfrow = c(2, 3))
```


# Plot mpg vs wt with regression line

```{r}
plot(mpg ~ wt, data = mtcars, col = "red", main = "MPG vs Weight", xlab = "Weight (1000 lbs)", ylab = "MPG")
abline(clm1, col = "pink")
```


# Plot mpg vs cyl with regression line

```{r}

plot(mpg ~ cyl, data = mtcars, col = "blue", main = "MPG vs Cylinders", xlab = "Cylinders", ylab = "MPG")
abline(clm2, col = "lightblue")
```



# Plot mpg vs hp with regression line

```{r}
plot(mpg ~ hp, data = mtcars, col = "seagreen", main = "MPG vs Horsepower", xlab = "Horsepower", ylab = "MPG")
abline(clm3, col = "green")
```


```{r}
summary(clm1)
summary(clm2)
summary(clm3)
```


```{r}
install.packages("gridExtra")
library(gridExtra)
```



# Load the mtcars dataset
```{r}
data(mtcars)
```


# Correlation
```{r}
cor(Orange$circumference,Orange$age)
```



# Create linear models

```{r}
clm1 <- lm(mpg ~ wt, data = mtcars)
clm2 <- lm(mpg ~ cyl, data = mtcars)
clm3 <- lm(mpg ~ hp, data = mtcars)


plot(clm1, which = 1:3)
plot(clm2, which = 1:3)
plot(clm3, which = 1:3)


sum((mtcars$mpg - clm1$fit)^2 )
sum((clm1$fit - mean(mtcars$mpg))^2 )
sum((mtcars$mpg - mean(mtcars$mpg))^2 )

sum((mtcars$mpg - clm2$fit)^2 )
sum((clm2$fit - mean(mtcars$mpg))^2 )
sum((mtcars$mpg - mean(mtcars$mpg))^2 )

sum((mtcars$mpg - clm3$fit)^2 )
sum((clm3$fit - mean(mtcars$mpg))^2 )
sum((mtcars$mpg - mean(mtcars$mpg))^2 )


```



# Create the summary statistics using pander
```{r}
summary_clm1 <- pander_return(summary(clm1))
summary_clm2 <- pander_return(summary(clm2))
summary_clm3 <- pander_return(summary(clm3))

```


# Create the plots using ggplot2

```{r}

plot1 <- ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point(color = "red") +
  geom_smooth(method = "lm", se = FALSE, color = "pink") +
  ggtitle("MPG vs Weight") +
  xlab("Weight (1000 lbs)") +
  ylab("MPG") + 
  theme_bw()

plot2 <- ggplot(mtcars, aes(x = cyl, y = mpg)) +
  geom_point(color = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "lightblue") +
  ggtitle("MPG vs Cylinders") +
  xlab("Cylinders") +
  ylab("MPG") + 
  theme_bw()

plot3 <- ggplot(mtcars, aes(x = hp, y = mpg)) +
  geom_point(color = "seagreen") +
  geom_smooth(method = "lm", se = FALSE, color = "green") +
  ggtitle("MPG vs Horsepower") +
  xlab("Horsepower") +
  ylab("MPG") + 
  theme_bw()
```



# Combine the plots and summaries using gridExtra


```{r}
grid.arrange(
  plot1, plot2, plot3,
  tableGrob(summary_clm1),
  tableGrob(summary_clm2),
  tableGrob(summary_clm3),
  ncol = 3
)
```



# Skills Quiz

```{r}
v <- c(-0.266, 0.489, -0.486, 0.569, -0.306)
r2 <- sum(v^2)
r2


```
```{r}
y_values <- c(3.78, 6.08, 6.65, 9.25, 9.92)
residuals <- c(-0.266, 0.489, -0.486, 0.569, -0.306)

# Calculate mean of y-values
y_mean <- mean(y_values)

# Calculate SS_tot
ss_tot <- sum((y_values - y_mean)^2)

# Calculate SS_res
ss_res <- sum(residuals^2)

# Calculate R-squared
r_squared <- 1 - (ss_res / ss_tot)

# Print R-squared
r_squared
```



```{r}
mtc.lm <- lm(disp ~ wt, data=mtcars)
summary(mtc.lm)
```
```{r}
mtc.lm <- lm(mpg ~ wt, data=mtcars)
pred <- predict(mtc.lm, newdata=data.frame(wt=2.7))
pred
```
```{r}
pred-21
```



