---
title: "Skills Quiz: Different Types of Models"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
    toc: true
    toc_float: true
---


## Instructions

Use this file to keep a record of your work as you complete the "Skills Quiz: Different Types of Models" assignment in Canvas.


```{r, message=FALSE, warning=FALSE}
pacman::p_load(tidyverse,dplyr, mosaicData)
```


----

<!-- Note: The {} after each Problem and Part header allows you to keep track of what work you have completed. Write something like {Done} once you complete each problem and your html file will then show you a nice summary of what you have "done" already. -->



## Problem 1 {}

Consider the scatterplot shown here of a single residence's monthly gas bill according to the month of the year. See `?Utilities` for more details on the data.

```{r}

gb <- lm(gasbill ~ month + I(month^2), data=Utilities)
β <- coef(gb)

# Plot the original data points
plot(gasbill ~ month, data=Utilities, main="Single Residence in Minnesota", 
     xlab="Month of the Year", ylab="Monthly Gas Bill (US Dollars)")

# Add the quadratic curve using the curve function
curve(β[1] + β[2] * x + β[3] * x^2, add=TRUE, col="red")

summary(gb)

```





### Part (a)

Add the estimated regression function to the scatterplot above and state the function below.

<div class="YourAnswer">

$$
  Y = 284 - 76 * x + 5.4 x^2
$$

</div>

### Part (b)

Diagnose the appropriateness of this regression model. How well does it fit the data? 

<span style="font-size:.8em;">Be sure to provide diagnostic plots and supporting arguments for your claims.</span>

<div class="YourAnswer">

Questionably

```{r}
View(Utilities)
plot(gb,which=1:3)
```

</div>


### Part (c)

What range of possible gas bill amounts do you predict for the September bill? How confident are you in your prediction?

<div class="YourAnswer">

```{r}
pred_conf_int <- predict(gb, newdata = data.frame(month = 9),
                         interval = "prediction", level = 0.95)

# Display the confidence interval

pred_conf_int
```

Type your answer here...

</div>


### Part (d)

Compute the mean of just September's gas bills and draw a horizontal reference line for this mean on your scatterplot. How does this mean compare to your prediction in Part (c)?

<div class="YourAnswer">

```{r}
plot(gasbill ~ month, data=Utilities, main="Single Residence in Minnesota", 
     xlab="Month of the Year", ylab="Monthly Gas Bill (US Dollars)")

abline(h=mean(Utilities$gasbill),col="red")
```

```{r}
sep_gb <- Utilities %>% filter(month==9)
sep_pred <- predict(gb, newdata = data.frame(month = 9))

print(mean(sep_gb$gasbill))

# Difference in mean vs prediction
print( mean(sep_gb$gasbill) - sep_pred)
```


Type your answer here...

</div>




### Part (e)

State the values of $R^2$, $MSE$, and the residual standard error. What do each of these values tell us about the regression model?

<div class="YourAnswer">

```{r}
summary(gb)
```
```{r}
model = gb
rss <- sum(residuals(model)^2)
# df = n - 2, df is already - 2, lose df for more parameters
df <- df.residual(model)
mse <- rss / df
mse
```

Type your answer here...

</div>

<br />



## Problem 2

View the `mtcars` dataset and corresponding help file `?mtcars`.

Perform a regression that predicts the miles per gallon `mpg` of the vehicle based on the quarter mile time `qsec` and transmission type `am` of the vehicle.


### Part (a)

Plot the data and your fitted regression model.

<div class="YourAnswer">

```{r}
View(mtcars)
```

```{r}
mtc.lm <- lm(mpg ~ qsec + am + qsec:am,data=mtcars)
summary(mtc.lm)
```

```{r}
# alt 225 for beta
ß <- coef(mtc.lm)

y1 <- function(x) ß[1] + ß[2] * x + ß[3] * 0 + ß[4] * x * 0
y2 <- function(x) ß[1] + ß[2] * x + ß[3] * 1 + ß[4] * x * 1

# Plotting
ggplot(mtcars, aes(x = qsec, y = mpg, color = factor(am))) +
  geom_point() +
  stat_function(fun = y1, aes(color = "Automatic")) +
  stat_function(fun = y2, aes(color = "Manual")) +
  #scale_color_manual(values = c("blue", "red"), labels = c("Automatic", "Manual")) +
  labs(x = "Quarter Mile Time (qsec)", y = "Miles per Gallon (mpg)") +
  theme_bw()
```


</div>


### Part (b)

State the fitted regression model.

<div class="YourAnswer">

$$
 \hat{Y} = -9 + 1.4*X_1 + -14.5 * X_2 + 1.3 * X_1*X_2 + \epsilon_i
$$

</div>

### Part (c)

Use `summary(...)` to perform an appropriate t test to determine if the interaction term is needed in this regression model.

<div class="YourAnswer">

```{r}

```

Type your answer here...

</div>


### Part (d)

Diagnose the appropriateness of this regression model. How well does it fit the data? 

<span style="font-size:.8em;">Be sure to provide diagnostic plots and supporting arguments for your claims.</span>

<div class="YourAnswer">

```{r}
plot(mtc.lm,which=1:3)
```

Type your answer here...

</div>


### Part (e)

State the $R^2$ and residual standard error values for this model. What do these values show?

<div class="YourAnswer">

```{r}
.722 #r^2
3.343 # RSE
```

Type your answer here...

</div>

<br />



## Problem 3

View the `mtcars` dataset and corresponding help file `?mtcars`.

Perform a regression that predicts the quarter mile time `qsec` of the vehicle based on the displacement of the engine `disp` and transmission type `am` of the vehicle according to the model:

$$
  \underbrace{Y_i}_\text{qsec} = \beta_0 + \beta_1 \underbrace{X_{1i}}_\text{disp} + \beta_2 X_{1i}^2 + \beta_3 \underbrace{X_{2i}}_\text{am == 1} + \beta_4 X_{1i}X_{2i} + \beta_5 X_{1i}^2X_{2i} + \epsilon_i
$$

where $\epsilon_i \sim N(0, \sigma^2)$.

### Part (a)

Plot the data and your fitted regression model.

<div class="YourAnswer">

```{r}
  mtc2.lm <- lm(qsec ~ disp + am + disp:am + I(disp^2) + I(disp^2):am, data=mtcars)
  summary(mtc2.lm)
```

```{r}
ß <- coef(mtc2.lm)

# Define functions for predicted values
y1 <- function(x) ß[1] + ß[2] * x + ß[3] * 0 + ß[4] * x^2 + ß[5] * x * 0 + ß[6] * x^2 * 0
y2 <- function(x) ß[1] + ß[2] * x + ß[3] * 1 + ß[4] * x^2 + ß[5] * x * 1 + ß[6] * x^2 * 1

# Plotting
ggplot(mtcars, aes(x = disp, y = qsec, color = factor(am))) +
  geom_point() +
  stat_function(fun = y1, aes(color = "Automatic")) +
  stat_function(fun = y2, aes(color = "Manual")) +
  #   scale_color_manual(values = c("blue", "red"), labels = c("Automatic", "Manual")) +
  labs(x = "Engine Displacement (disp)", y = "Quarter Mile Time (qsec)") +
  theme_bw()
```

Type your answer here...

</div>

### Part (b)

State the fitted regression model.

<div class="YourAnswer">

$$
  \hat{Y}_i = ...
$$

</div>


### Part (c)

Use `summary(...)` to perform appropriate t tests to determine which interaction terms are needed in this regression model.

<div class="YourAnswer">

```{r}
#Type your code here...
```

</div>

### Part (d)

Diagnose the appropriateness of this regression model. How well does it fit the data? 

<span style="font-size:.8em;">Be sure to provide diagnostic plots and supporting arguments for your claims.</span>


<div class="YourAnswer">

```{r}
plot(mtc2.lm,which=1:3)
```

Type your answer here...

</div>

### Part (e) 

Drop the "least significant" term from the model. This is a silly statement really, but is an accepted practice when searching for a "best" model in regression analysis. What has changed?

<div class="YourAnswer">

```{r}
mtc2.lm <- lm(qsec ~ disp + am + I(disp^2) + I(disp^2):am, data=mtcars)
summary(mtc2.lm)
```

Type your answer here...

</div>


## Q4

```{r}
ß1 <- -2
ß2 <- 3
ß3 <- 4
ß4 <- 4
ß5 <- 2
ß6 <- -7

x2 <- sample(c(0,1),100,replace=TRUE)

# Define the functions
fun <- function(x) ß1 + ß2 * x + ß3 * x2 + ß4 * x^2 + ß5 * x2 + ß6 * x^2 * x2

sigma <- .001 #change to whatever positive number you want
epsilon_i <- rnorm(n, 0, sigma)


# Generate data
set.seed(123)  # Setting seed for reproducibility
n <- 100
x <- runif(n, -2, 2)
y = fun(x) + epsilon_i

plotdata <- data.frame(x, y, x2)

tq.lm <- lm(y ~ x + x2 + x:x2 + I(x^2) + I(x^2):x2, data=plotdata)
ß <- coef(tq.lm)

y1 <- function(x) ß[1] + ß[2] * x + ß[3] * 0 + ß[4] * x^2 + ß[5] * 0 + ß[6] * x^2 * 0
y2 <- function(x) ß[1] + ß[2] * x + ß[3] * 1 + ß[4] * x^2 + ß[5] * 1 + ß[6] * x^2 * 1

# Plotting
ggplot(plotdata, aes(x = x, y = y)) +
  geom_point() +
  labs(x = "X", y = "Y", title = "Generated Data Plot") +
  stat_function(fun = y1, aes(color = "Automatic")) +
  stat_function(fun = y2, aes(color = "Manual")) +
  ylim(-10, 10) +
  # scale_color_manual(values = c("blue", "red"), labels = c("Function 1", "Function 2")) +
  theme_bw()
```

## QUIZ

### Q1
```{r}
library(Ecdat)
```


## Code to make this graphic...

> ggplot(Wages1, aes(x=school, y=wage)) + geom_point(color=rgb(.3,.3,.2,.4)) + geom_smooth(method="lm", formula= y ~ poly(x, 2, raw=TRUE), se=F, col=rgb(.3,.3,.2,.8)) + theme_bw() + labs(title="Best wages for highest schooling... or no schooling?", subtitle="Wages1 data, library(Ecdat)", x="Number of years of Schooling (school)", y="Current Hourly Wage in 1980's USD (wage)")

 

## Code needed to obtain the solution...

> summary(lm(wage ~ school + I(school^2), data=Wages1))

Call:
lm(formula = wage ~ school + I(school^2), data = Wages1)

Residuals:
Min 1Q Median 3Q Max
-7.337 -1.998 -0.474 1.444 34.554

Coefficients:
                      Estimate    Std. Error     t value    Pr(>|t|)
(Intercept)       6.13334      1.46274       4.193    2.82e-05 ***
school            -0.68354      0.25743      -2.655    0.00796 **
I(school^2)      0.05488       0.01129       4.859   1.23e-06 ***
---
Signif. codes: 0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Residual standard error: 3.126 on 3291 degrees of freedom
Multiple R-squared: 0.08636, Adjusted R-squared: 0.0858
F-statistic: 155.5 on 2 and 3291 DF, p-value: < 2.2e-16

 

 

## Code to draw the plot in base graphics...

> plot(wage ~ school, data=Wages1, pch=16, col=rgb(.3,.3,.2,.4), main="Best wages for highest schooling... or no schooling?",  xlab="Number of years of Schooling (school)", ylab="Current Hourly Wage in 1980's USD (wage)")

> mtext(side=3, text="Wages1 data, library(Ecdat)")

> abline(v=seq(2,16,2), h=seq(0,40,10), lty=2, col=rgb(.2,.2,.2,.2))

> wages.lm <- lm(wage ~ school + I(school^2), data=Wages1)

> b <- coef(wages.lm)

> curve(b[1] + b[2]*x + b[3]*x^2, add=TRUE, col=rgb(.3,.3,.2,.8), lwd=2)

```{r}
View(Wages1)
wages.lm <- lm(wage ~ school + I(school^2),data=Wages1)
summary(wages.lm)
```
### Q2
it was groups 1 and 2, so it only added the presence of the intercept change.

Since the equation only allows the y-intercept to be changed by the term, and there is no interaction term allowing the slope to change, then the correct answer is the plot that has lines with equal slopes. In other words, this model forces parallel lines. Only models that include an interaction term can have differing slopes.

### Q3

```{r}
View(Loblolly)
?Loblolly
```

```{r}
ll.lm1 <- lm(height ~ age + I(age^2), data=Loblolly)
ll.lm2 <- lm(height ~ age + I(age^2) + I(age^3),data=Loblolly)
ll.lm3 <- lm(height ~ age + I(age^2) + I(age^3) + I(age^4),data=Loblolly)
ll.lm4 <- lm(height ~ age + I(age^2) + I(age^3) + I(age^4) + I(age^5),data=Loblolly)
```

```{r}
plot(ll.lm1, which=1:3)
```

```{r}
plot(ll.lm2, which=1:3)
```

```{r}
plot(ll.lm3, which=1:3)
```

```{r}
plot(ll.lm4, which=1:3)
```

Running the codes:

plot(lm(height ~ age + I(age^2), data=Loblolly), which=1)

plot(lm(height ~ age + I(age^2) + I(age^3), data=Loblolly), which=1)

plot(lm(height ~ age + I(age^2) + I(age^3) + I(age^4), data=Loblolly), which=1)

plot(lm(height ~ age + I(age^2) + I(age^3) + I(age^4) + I(age^5), data=Loblolly), which=1)

will show the last one has the "flattest" red line, witnessing the greatest "linear" fit of the data. If you wanted to resolve the lack of constant variance, you could then apply: 

boxCox(lm(height ~ age + I(age^2) + I(age^3) + I(age^4) + I(age^5), data=Loblolly))

and the resulting transformation looks quite nice:

plot(lm(sqrt(sqrt(height)) ~ age + I(age^2) + I(age^3) + I(age^4) + I(age^5), data=Loblolly), which=1)

except linearity is still lacking...

plot(lm(sqrt(sqrt(height)) ~ age + I(age^2) + I(age^3) + I(age^4) + I(age^5), data=Loblolly), which=2)


<style>

.YourAnswer {
  color: #317eac;
  padding: 10px;
  border-style: solid;
  border-width: 2px;
  border-color: skyblue4;
  border-radius: 5px;
}

</style>

 
 