---
title: "Recovering a True Model"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

```{r, message=FALSE, warning=FALSE}
rbdata_beg <- read.csv("rbdata_beginner-1.csv", header=TRUE, stringsAsFactors = TRUE)
rbdata_int <- read.csv("rbdata_intermediate.csv", header=TRUE, stringsAsFactors = TRUE)
rbdata_adv <- read.csv("rbdata_advanced.csv", header=TRUE, stringsAsFactors = TRUE)
```

## Part 1 - Recovering a True Model (Beginner Level)

We believe the true model to be...

$$
  Y_i = \underbrace{B}_\text{The True Model} + \epsilon_i
$$

And estimate that model by...

$$
  \hat{Y}_i = 92.34 + -8.57 X_{3i} + 6.1 X_{5i} + .099 X_{3i} X_{5i}
$$
```{r}
View(rdata_beg)
pairs(rbdata_beg, panel=panel.smooth, col=as.factor(rbdata_beg$X3))

lm1 <- lm(Y ~ X3*X5, data=rbdata_beg)
summary(lm1)
```

with our estimate of $\sigma$ as ...

```{r}

```


## Part 2 - Recovering a True Model (Intermediate Level)

We believe the true model to be...

$$
  Y_i = \underbrace{...}_\text{The True Model} + \epsilon_i
$$

And estimate that model by...

$$
  \hat{Y}_i = ...
$$

with our estimate of $\sigma$ as ...

```{r}
View(rbdata_int)
pairs(rbdata_int, col = as.factor(rbdata_int$x2))
pairs(rbdata_int, col = as.factor(rbdata_int$x3))
```



## Part 3 - Recovering a True Model (Advanced Level)

We believe the true model to be...

$$
  Y_i = \underbrace{...}_\text{The True Model} + \epsilon_i
$$

And estimate that model by...

$$
  \hat{Y}_i = ...
$$

with our estimate of $\sigma$ as ...


```{r}
# Use this on the intermediate data

ggplot(rbdata_int, aes(x=x2,y=y,color=interaction(x4,x5,x8)))+
  geom_point() +
  # see the first chart, how x4 and 5 and 5 and x8 can't activate simultaneously
  geom_point(aes(y=lm2$fit),cex=0.5) +
  facet_wrap(~interaction(x4,x5,x8)) +
  theme_bw()
```


```{r}
View(rdata_adv)
pairs(rbdata_adv, panel=panel.smooth, col=as.factor(rbdata_adv$x4))
pairs(rbdata_adv, panel=panel.smooth, col=as.factor(rbdata_adv$x5))
pairs(rbdata_adv, panel=panel.smooth, col=as.factor(rbdata_adv$x8))
```
```{r}
ggplot(rbdata_adv, aes(x=x2,y=y,color=interaction(x4,x5,x8)))+
  geom_point() +
  # see the first chart, how x4 and 5 and 5 and x8 can't activate simultaneously
  facet_wrap(~interaction(x4,x5,x8)) +
  theme_bw()

# Using terms that seem like they work, lets just see what happens
lm1 <- lm(y ~ x2*x4*x5*x8 *I(x2^2), data=rbdata_adv)
summary(lm1)
# NAs - you've used some things that can't happen

# Let's construct an lm with what we think is actually effecting stuff
lm2 <- lm(y~ x2 + I(x2^2) +
            x4 + x4:x2 + x4:I(x2^2) +
            x5 + x5:x2 + x5:I(x2^2) +
            x8 + x8:x2 + x8:I(x2^2) + x8:I(x2^3) + x8:I(x2^4) +
            x4:x8 + x4:x8:x2 + x4:x8:I(x2^2) + x4:x8:I(x2^3) + x4:x8:I(x2^4),
            data=rbdata_adv)

summary(lm2)

ggplot(rbdata_adv, aes(x=x2,y=y,color=interaction(x4,x5,x8)))+
  geom_point() +
  # see the first chart, how x4 and 5 and 5 and x8 can't activate simultaneously
  geom_point(aes(y=lm2$fit),cex=0.5) +
  facet_wrap(~interaction(x4,x5,x8)) +
  theme_bw()

```


Note x4 and x5. They are missing a dot in the top right corner. Also x5 and x8. These are binary variables. Slope is exciting. Note y - x8,

Note y-x2 in x4

```{r}
library(plotly)
library(reshape2)
```

```{r}
helix1 <- filter(rbdata_adv,x4 == 0, x5 == 0, x8 == 1)
helix1$side = 0
helix2 <- filter(rbdata_adv,x4 == 1, x5 == 0, x8 == 1)
helix1$side = 1
```


```{r}

helix_data = merge(helix1, helix2)

#Graph Resolution (more important for more complex shapes)
graph_reso <- 0.5

#Setup Axis
axis_x <- seq(min(airquality$Temp), max(airquality$Temp), by = graph_reso)
axis_y <- seq(min(airquality$Month), max(airquality$Month), by = graph_reso)

#Sample points
air_surface <- expand.grid(Temp = axis_x, Month = axis_y, KEEP.OUT.ATTRS=F)
air_surface$Z <- predict.lm(air_lm, newdata = air_surface)
air_surface <- acast(air_surface, Month ~ Temp, value.var = "Z") #y ~ x

#Create scatterplot
plot_ly(rbdata_adv, 
        x = ~x2, 
        y = ~y, 
        z = ~Ozone,
        text = rownames(airquality), 
        type = "scatter3d", 
        mode = "markers") %>%
  add_trace(z = air_surface,
            x = axis_x,
            y = axis_y,
            type = "surface")
```

